---
title: "Introduction to avg_word_len"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{avg_word_len}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

During Natural Language Processing (NLP), data scientists may want:
- extract additional characteristics from text data
- engineer their own features using these additional characteristics
- train their models with their engineered features for better model results

The `avg_word_len` function from the `textfeatureinfo` package is one of four functions that aids in this text data extraction process. Specifically, `avg_word_len` makes it easy to extract the average length of words inside of a piece of text. 

This document provides detailed instructions on how to apply the `avg_word_len` function on text data for quick access to a new feature that can be added into various machine learning models.

You can start by loading the `textfeatureinfo` package:
```{r setup}
library(textfeatureinfor)
```



