---
title: "Introduction to avg_word_len"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{avg_word_len}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

During Natural Language Processing (NLP), data scientists may want:
- extract additional characteristics from text data
- engineer their own features using these additional characteristics
- train their models with their engineered features for better model results

The `avg_word_len` function from the `textfeatureinfo` package is one of four functions that aids in this text data extraction process. Specifically, `avg_word_len` makes it easy to extract the average length of words inside of a piece of text. 

This document provides detailed instructions on how to apply the `avg_word_len` function on text data for quick access to a new feature that can be added into various machine learning models.

You can start by loading the `textfeatureinfo` package:
```{r setup}
library(textfeatureinfor)
```

## Quick example for one piece of text

To understand the basics of the `avg_word_len` function, we can take a look at it's interaction with a simple text object. 

```{r}
text <- "Star wars is my favourite movie"

round(avg_word_len(text), 2) # the output of avg_word_len is rounded to 2 decimals
```

## Example with a dataframe

Now that we understand how to use this function, let's apply `avg_word_len` to an entire data set. This is more practical aspect of this function considering that feature extraction will need to be done for entire dataframes in order to be useful for machine learning models.

```{r}


```

