---
title: "Introduction to avg_word_len"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{avg_word_len}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Background


During Natural Language Processing (NLP), data scientists may want:
- extract additional characteristics from text data
- engineer their own features using these additional characteristics
- train their models with their engineered features for better model results

The `avg_word_len` function from the `textfeatureinfo` package is one of four functions that aids in this text data extraction process. Specifically, `avg_word_len` makes it easy to extract the average length of words inside of a piece of text. 

This document provides detailed instructions on how to apply the `avg_word_len` function on text data for quick access to a new feature that can be added into various machine learning models.

You can start by loading the `textfeatureinfo` package:
```{r setup}
library(textfeatureinfor)
```

## Quick example with one piece of text


To understand the basics of the `avg_word_len` function, we can take a look at its interaction with a simple text object. 

```{r}
text <- "Star wars is my favourite movie"

round(avg_word_len(text), 2) # the output of avg_word_len is rounded to 2 decimals
```

From this basic example, we can see that the average length of words in this piece of text is `4.33`.

## Example with a dataframe


Now that we understand how to use this function, let's apply `avg_word_len` to an entire data set. This is more practical aspect of this function considering that feature extraction will need to be done for entire data-frames in order to be useful for machine learning models.

Let's first create a toy data-set with 5 examples:
```{r}
toy_data <- data.frame(messages = c(
  "Hello World",
  "Star wars is my favourite movie",
  "Never have I ever",
  "What a wonderful world",
  "You should watch Spiderman: no way home"
  ))
toy_data
```

Now, we will leverage the `avg_word_len` function to extract the average word lengths from the toy data and add it as a new column in the toy data frame

```{r}
# applies avg_word_len function to each row in the data frame
toy_data["average_word_lengths"] <- apply(toy_data["messages"], 
                                          1,
                                          avg_word_len) 

toy_data["average_word_lengths"] <- round(toy_data["average_word_lengths"], 2)
toy_data
```

Now we have access to the average length of the words for each example message in the toy data frame.

## Summary


The `avg_word_len` function is simple, but has broad implications. It allows for easier extraction of useful text information in one line of code. This feature is useful for natural language processing and may prove useful when building machine learning models.

