---
title: "Introduction to avg_word_len"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{avg_word_len}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Background
___

During Natural Language Processing (NLP), data scientists may want:
- extract additional characteristics from text data
- engineer their own features using these additional characteristics
- train their models with their engineered features for better model results

The `avg_word_len` function from the `textfeatureinfo` package is one of four functions that aids in this text data extraction process. Specifically, `avg_word_len` makes it easy to extract the average length of words inside of a piece of text. 

This document provides detailed instructions on how to apply the `avg_word_len` function on text data for quick access to a new feature that can be added into various machine learning models.

You can start by loading the `textfeatureinfo` package:
```{r setup}
library(textfeatureinfor)
```

## Quick example for one piece of text
___

To understand the basics of the `avg_word_len` function, we can take a look at it's interaction with a simple text object. 

```{r}
text <- "Star wars is my favourite movie"

round(avg_word_len(text), 2) # the output of avg_word_len is rounded to 2 decimals
```

## Example with a dataframe
___

Now that we understand how to use this function, let's apply `avg_word_len` to an entire data set. This is more practical aspect of this function considering that feature extraction will need to be done for entire data-frames in order to be useful for machine learning models.

Let's first create a toy data-set with 5 examples:
```{r}
toy_data <- data.frame(messages = c(
  "Hello World",
  "Star wars is my favourite movie",
  "Never have I ever",
  "What a wonderful world",
  "You should watch Spiderman: no way home"
  ))
toy_data
```

Now, we will leverage the `avg_word_len` function to extract the average word lengths from the toy data and add it as a new column in the toy data frame

```{r}
# applies avg_word_len function to each row in the data frame
toy_data["average_word_lengths"] <- apply(toy_data["messages"], 
                                          1,
                                          avg_word_len) 

toy_data["average_word_lengths"] <- round(toy_data["average_word_lengths"], 2)
toy_data
```


